---
title: "R Notebook"
output: html_notebook
---

## Part 1

What is our goal? We want to model senlity (s)(binary) with a logistic regression against WAIS score (x). We've grouped the observations with 'summaryBy' because we have some observations with the same WAIS score.

```{r}
library(dobson)
library(doBy)
sen.grp <- summaryBy(s~x, data=senility, FUN=c(sum, length))
names(sen.grp) <- c("x","y","n")
```

The 'cbind' function creates a 2*n matrix containing the number of senility counts in the first column for a WAIS score, and the non-senility count in the second column. Not sure why.
```{r}
cbind(sen.grp$y, sen.grp$n - sen.grp$y)
```

Next we set up the generalised linear model of the matrix as described above as a function of x, the WAIS score.
```{r}
senility.grp <- glm(cbind(y, n - y) ~ x, data=sen.grp, family=binomial)
```

Then we have a look at the high level summary of the model.
```{r}
summary(senility.grp)
```

------

## Part 2

Now we want to test the fit of the model, by looking at the Pearson Chi^2 and Deviance statistics

For calculating the Chi^2, we first calculate the critical value:
```{r}
crit.val <- qchisq(0.95, df=nrow(sen.grp)-2)
```

Then we can do a little programming to test the residuals
```{r}
P.res <- residuals(senility.grp, type = "pearson")
if(sum(P.res^2) > crit.val){
  cat("We reject the null hypothesis that the model is correct")
}else{
  cat("We cannot reject the null hypothesis that the model is correct")
}
```

No luck so far, so we try the Deviance statistic

```{r}
D.res <- residuals(senility.grp, type = "deviance")
if(sum(D.res^2) > crit.val){
  cat("We reject the null hypothesis that the model is correct")
}else{
  cat("We cannot reject the null hypothesis that the model is correct")
}
```

Bad news by the looks.

Next we reproduce a table from the textbook for some reason...
```{r}
cbind(sen.grp, senility.grp$fitted.values, P.res, D.res)
```

Now we look at the standardised residuals (Pearson and Deviance) as function of the WAIS score.
```{r}
stanresPearson <- P.res/sqrt(1-hatvalues(senility.grp))
stanresDeviance <- D.res/sqrt(1-hatvalues(senility.grp))

# Alternative for diagonal elements of hat matrix:
# influence(senility.grp)$hat

plot(sen.grp$x,stanresDeviance, ylab="Standardized Deviance Residuals",
     xlab="WAIS score", ylim=c(-2,2))
plot(sen.grp$x,stanresPearson, ylab="Standardized Pearson Residuals",
     xlab="WAIS score", ylim=c(-2,2))
```

Looks like the residuals are the same for both of the tests, and that there are a couple of potential outliers? No answer given.

Now we try the Hosmer-Lemeshow statistic with the use of the 'ResourceSelection' library in R.
```{r}
library(ResourceSelection)
res.glm <- glm(s~x, data=senility, family="binomial")
h <- hoslem.test(senility$s,fitted(res.glm),g=3)
h$expected
```

Here are some more statistics
```{r}
h$observed
```

```{r}
h$statistic
```

```{r}
h$p.value
```

"The p-value is above $\alpha = 0.05$, so at the 5% level of significance the null hypothesis that the observed and expected proportions are the same across all doses cannot be rejected (evidence of a good fit from the model)."
