---
title: "Generalised Addititve Models"
output: html_notebook
---

# Activity 1
Perform a series of ANOVA tests in order to determine
which of these three models is best for predicting wage:
1) a GAM that excludes 'year';
2) a GAM that uses a linear function of 'year';
3) a GAM that uses a spline function of 'year'.

```{r}
library(ISLR)
library(splines)
library(gam)

data('Wage')
attach(Wage)

gam1 = gam(wage ~ ns(age, 5) + education, data=Wage)
gam2 = gam(wage ~ ns(age, 5) + education + year, data=Wage)
gam3 = gam(wage ~ ns(age, 5) + education + ns(year, 4), data=Wage)

anova(gam1, gam2, gam3, test='F')
```

"We find that there is compelling evidence that a GAM with
a linear function of year is bettern than a GAM that does 
not include year at all. However, there is no evidence
that a non-linear function of year is needed."

------

# Activity 2
Analyse 'out-of-state' data in the College data set from ISLR.
0) Split the data into training, validation, and testing sets.
1) Perform stepwise selection on the training set in order to
identify a satisfactory model that uses just a subset of the
predictors.
2) Fit a GAM on the training data. Plot the results and discuss
findings.
3) Evaluate the model obtained on the test set and discuss the
results.
4) For which variables (if any) is there evidence of a non-linear
relationship with the response.

# 0
```{r, message=FALSE}
library(ISLR)
library(leaps)
library(glmnet)
library(gam)

attach(College)

set.seed(0)

# 'Split' the data into testing, validation, and training sets:
# - Create a vector of College table length with numbers 1 through 3.
# - Probability of a 1 is 0.5, for 2 is 0.25, and for 3 is 0.25.
# - Index of 1 will be used for testing, 2 for validation, 3 for training.
data_index = sample(1:3, nrow(College), replace=T, prob=c(0.5,0.25,0.25))
```

# 1
First we set up the regression subsets on the training data.
```{r}
p = ncol(College)-1 # Number of predictor columns
fit_forward = regsubsets(
  Outstate ~ ., data=College[data_index==1, ],
  nvmax=p, method='forward'
)

reg_summary = summary(fit_forward)
#reg_summary
```

Then we use the validation data to compute MSE to determine
the optimal parameter subset.
```{r}
validation_matrix = model.matrix(Outstate ~ ., data=College[data_index==2, ]) # ?
validation_errors = rep(NA, p) # Blank column of NAs at length of parameter vector

for(i in 1:p){ # ?
  coeff = coef(fit_forward, id=i)
  pred = validation_matrix[, names(coeff)] %*% coeff
  validation_errors[i] = mean((College$Outstate[data_index==2] - pred)^2)
}

k = which.min(validation_errors) # Optimal # of parameters based on MSE

plot(
  validation_errors, xlab='Number of variables', ylab='Validation MSE',
  pch=19, type='b'
)

abline(v=k, col='red') # Vertical line for # of parameter at min MSE

grid()
```

The best model appears to be one with 9 parameters.
But we decide 'for simplicity' to pick a model with only 3.
```{r}
k = 3
coeff = coef(fit_forward, id=k)

test_matrix = model.matrix(Outstate ~ ., data=College[data_index==3, ])
pred = test_matrix[, names(coeff)] %*% coeff
test_error = mean((College$Outstate[data_index==3] - pred)^2)
coeff
test_error # MSE for 3 parameter
```


# 2
Now we fit a GAM using the 3 predictors found above.
We see that both Expend and Room.Board seem to be non-linear.
```{r}
data_index[data_index==2] = 1 # Use 75% data for training?
data_index[data_index==3] = 2 # And 25% data for testing?

fit_gam = gam(
  Outstate ~ s(Expend, 4) + s(Room.Board, 4) + Private,
  data=College[data_index==1, ]
)

par(mfrow=c(1,3))
plot(fit_gam, se=TRUE, col='blue')
```

# 3
Now we evaluate the model on the test set.
```{r}
y_hat = predict(fit_gam, newdata=College[data_index==2, ])
MSE = mean((College[data_index==2, ]$Outstate - y_hat)^2)
MSE
```

The MSE for the GAM model is lower than the 3 parameter model
put forward by the (linear) regression subsets.

# 4
The output below shows that both Expend and Room.Board seem
to be non-linear (Private is boolean).
```{r}
summary(fit_gam)
```

